### 2014-11-22 docs/ms.Rmd
Integrate Kate's latest edits to ms.rmd

## Abstract
The colours of plants and animals have long fascinated biological scientists, but it is only recently that quantitative colour measurement techniques have become readily available. Spectrometry and digital photography are now relatively accessible and common techniques for the appraisal of biological colour, however, their rapid uptake has not been followed up by clear guidelines on which methodological details are crucial to report. We identified a number of details regarding measurement and analysis of colouration which have the potential to affect the outcomes of research. We reviewed reporting of relevant methodology in the 2013 literature on biological colour, and have outlined explicit recommendations for future research. In providing a constructive criticism of current practices, and identifying areas for improvement, we aim to better the accountability, transparency and reproducibility of colour research in the life sciences.

@@

## Introduction
Generations of biologists have endeavored to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891]. For these biologists, identifying best practices in measuring colour and colour pattern has been a great challenge. Over the past 150 years, biologists have developed and adapted techniques that are now widespread, including spectrometry, photography, and visual modelling.

The application of spectrometry to biological questions revolutionized the field [@dyck_determination_1966]. Before its widespread use, researchers assigned categorical colour descriptions, arbitrary rankings or scores [e.g. @aldrich_ecogeographic_1991; @bailey_latitudinal_1978; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g. @burtt_analysis_1986; @key_temperature_1954]. Although no doubt applied carefully, these techniques limited progress. For example, many researchers designed colour assessment methods specifically for their projects [e.g. @aldrich_ecogeographic_1991; @bailey_latitudinal_1978] rendering data incomparable across studies. This is especially problematic when different measurement techniques are used on the same species [e.g. the house finch _Carpodacus mexicanus_; @aldrich_origin_1978; @hill_female_1990]. Likewise, because human vision is inherently different to most other animals [@bennett_sexual_1994; @douglas_spectral_2014], its use can lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].

@@

Complementary to spectrometry, digital photography is increasingly being used to quantify biological colouration. High-resolution cameras are inexpensive,  and allow instantaneous, repeatable sampling of multiple colour patches [@mckay_use_2013]. The main challenge for gathering colour data from photographs involve ensuring standardization, such as linearization and consistent lighting [@stevens_using_2007]. Most digital cameras used for the measurement of biological colouration are consumer products designed for the human visual system. Thus hardware modifications for ultraviolet and infrared imaging are generally necessary [e.g. @stevens_colour_2013].

Expansion in the availability of objective methods for the measurement of colour has been matched by advances in theory and analysis. In particular, the development of sensory models has produced significant insights into the way in which colourful stimuli are perceived [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998]. Where the receptor sensitivities of an animal’s visual system are known, the physiological response of that animal’s visual system when viewing a given colour can be predicted.

Sensory models typically attempt to describe the subjective perception of colour or brightness information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g. @endler_comparing_2005; @vorobyev_receptor_1998]. These models enable researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to formulate and test hypotheses.

Although relatively easy to implement, and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and perceptually processed and these can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012]. For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; @vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e. raw versus log-transformed), the use of any colour constancy mechanism (e.g. von Kries), as well as the photoreceptor signal-to-noise ratios, their relative densities, and the type of noise predominating (i.e. neural and/or receptor noise). These data are only available for a small number of species, such as humans and honeybees, and so assumptions and extrapolations are necessary when models are applied to other species. Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies [@pike_preserving_2012].

Since the repeatability and transparency of science is paramount, it is imperative that we regularly and constructively reflect on our own practices [@alsheikh_public_2011; @sandve_ten_2013]. The potential for subjectivity in colour measurement and the mathematical opacity of visual models for the average biologist means that, as in any field, is it critical that methods are reported rigorously. Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011]. The aims of this paper were to (a) review methods reporting in papers published on colour in 2013 from 22 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods. Our review specifically focuses on papers that quantify the external coloration of a plant or animal. These may be purely descriptive quantifications of color, or they may be incorporated into models of animal vision. Our hope is that this paper will provide a useful set of guidelines to improve the reproducibility and transparency of studies on colour.

@@

We developed a list of metrics we deemed critical to be reported for reproducibility of colour methods (Table 1) and analysis (Table 2), loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006]. These included quantitative methods including spectrophotometric and photographic techniques [as in @kemp_ornamental_2008; @chiao_visualization_2009] but not qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009]. For visual modeling techniques we characterized a set of criteria required for reproducing the analysis (Table 2). Much of these criteria specifically targeted the necessary parameters required to be explicitly defined by the user during analysis. These are essential if such analyses are to be reproduced.

@@

To determine the current state of methods reporting in the field we searched for papers from 2013 in 22 top colour science journals (_Animal Behaviour_, _Behavioral Ecology_, _Journal of Experimental Biology_, _Biology Letters_, _Proceedings of the Royal Society B: Biological Sciences_, _Biological Journal of the Linnean Society_, _Ethology_, _The American Naturalist_, _Ecology Letters_, _Journal of Evolutionary Biology_, _American Journal of Botany_, _Ecology and Evolution_, _Oikos_, _Ecology_, _Journal of Ecology_, _Behavioral Ecology and Sociobiology_, _Current Zoology_, _Naturwissenschaften_, _New Phytologist_, _PLoS ONE_, _Evolution_, _Functional Ecology_). For each of the journals, we used the Boolean search terms 'colour*', 'color*' or 'spectra*' and searched for these in the title or abstract from each of the journal home pages. We read the titles and abstracts of each paper and included only those papers that explicitly used either a spectrophotometer or camera to quantify colour signals. We excluded papers that were: 1) review papers; 2) methodological papers; 3) papers quantifying colour pattern rather than characteristics of each individual colour patch; and 4) papers taking microspectrophotometric measurments of retinal absorbance. To review the large number of papers returned from the searches, journals were haphazardly divided up between the authors. In order to reduce the risk of observer bias in our scoring, papers were read and reassessed by a total of three authors. Any discrepancies between the assessor scores were discussed and agreed upon prior to analysis. We then tabulated what details of the methods were reported against our criteria. The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep (TODO: GIVE IT A DOI?). We have kept the papers used in our dataset annoymous to avoid accusing specific colleagues of inadvertant reporting. We feel that the issue of reporting is something attributable to the ambiguity and lack of guidelines in our field and not something specific to individual research groups.

Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis. Overall, `r specs+both` studies used a spectrometer (henceforth "spec") to measure colour, while `r cam+both` used a camera. Of these studies, `r both` quantified colour using both a spec and a camera. Reporting of the make and model of the spec and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 2)`% and cameras = `r round(camModel, digits = 2)`%), however, only `r round(CamLight, digits = 2)`%  and `r round(specLight, digits = 2)`% of camera and spec studies reported a light source. In addition, the number of reflectance spectra and pixels averaged were only reported in `r round(specAVG, digits = 2)`% and `r round(camAVG, digits = 2)`% of studies, respectively. Surprisingly, dark standards were only reported in `r round(drk_std, digits = 2)`% of the spec studies, while white standards were reported in `r round(white_stdspec, digits = 2)`%. Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(intTime, digits = 2)`%), or the angle of the optical probe relative to the surface (`r round(angle, digits = 2)`%) or distance (`r round(dist, digits = 2)`%) the probe was held from the measurement surface.

Analysis varied across studies. Overall `r analysisType[1]` papers analyzed  data by generating colourimetric variables [i.e. 'spectral' measures of hue, saturation, and/or brightness; @montgomerie_analyzing_2006] whereas `r analysisType[4]` used visual models, `r analysisType[2]` used multiple analysis types, and `r analysisType[3]` used other forms of analysis (e.g. purely statistical analyses, such as PCA's on reflectance spectra). Of the studies quantifying colour using spectral reflectance curves `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models and `r analysisTypeSpec[3]` used other forms of analysis. Studies using colourimetric measurements defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 2)`% of the studies.

@@

While some of the values reported above seem alarming, it is important to note that papers often made reference to previous work for various methods (`r round(ref_stud, digits = 2)`%), so these figures are slightly liberal given that referenced work may have comprehensively covered some of these criteria. Despite this, many of the studies that did not reference previous work were lacking in their reporting of many aspects of our criteria suggesting the need for more comprehensive reporting of data collection and analysis methods within such studies.

@@

Our review of recent colour literature highlights some important reporting gaps that will hinder the reproducibility of studies of biological colouration. While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly believe that researchers can do a better job of reporting details that are essential to both the outcome of their results and reproducibility of their work.  While all criteria are important, some deserve special attention. In particular, we found that colour scientists could do a better job reporting light sources, the numbers of pixels or spectra averaged, and the type of standards used (i.e. white and black reference samples). More importantly, the distance and angle of the optical probe relative to the measurment surface should be standard reporting for all studies using spectrophometry. Visual models were regularly used to analyse spectral data, something we welcome. However, there were many important aspects of the analysis that were seldom reported, including the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of neural noise used in receptor-noise models. Photoreceptor densities should also be reported where possible even if using standard avian photorecptor densities. These all have important consequences on downstream analysis and the reproducibility of colour signal quantification [e.g. @kemp_ornamental_2008; @santos_comparison_2007]. In addition, we recommend the presentation of spectral reflectance curves where possible; these figures are visually appealing and allow for the rapid assesment of the nature and quality of the data, independent of downstream processing. Surprisingly, we found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had a figure showing these curves. When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investigated explicitly outlined.

Our objective in this paper was not to highlight shortcomings in single studies, but to bring attantion to common oversights in reporting that can influence the reproducibility of colour research and cross-study comparisons. By no means do we wish to say that studies lacking important details are not of significant value. Indeed, even in our own work [e.g. @umbers_bright_2013] we have failed to report important details, which in retrospect can lead to confusion about how such data were collected and analysed. This review of reporting is aimed at providing constructive criticism; it is only through critical appraisal and collegial discussion that we can identify such gaps as areas for improvement, and constructively contribute to the growth of our field. Such assessments are particularly important in an age where accountability, transparency and reproducibility are of utmost importance.

We believe that part of the issue at hand is simply that many researchers seldom have clear guidelines to follow when undertaking research, and may overlook outlining important aspects of methodology and analysis. We hope our critical evaluation of recent literature has highlighted a need for more careful description of how colour signals have been measured and analyzed. We are optimistic that this may help establish uniform reporting guidelines for publication, and to continue the advance of an exciting era in colour research.

>>>>

## Abstract
Over the last 50 years or so, our enhanced capacity for quantitative colour measurement has accelerated understanding of the evolution of biological colouration. However, while spectrometry and digital photography are inexpensive, accessible and efficient, details of exactly how data are collected and analysed are not reported consistently in the literature. Similarly, estimates of animals’ visual capacity are widely applied to colour data, but details on assumptions and settings in analyses are not reported completely. To quantify the variability of methods reporting in studies of colouration, we reviewed papers published in 2013 in 22 leading journals. Using the search terms colour* OR color* OR spectra* in the title or abstract, we identified 60 papers in which colour had been measured and analysed. We extracted the methods and noted the details included. We found a great deal of variability in the level of detail in which colour methods is reported. In response we have outlined explicit recommendations on the important methodological details required to ensure studies are transparent and repeatable. We hope that this will help to standardize methods reporting in the field of biological colouration paving the way for direct comparison across studies.

@@

## Introduction
Generations of biologists have endeavored to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891]. For these biologists, identifying best practices in measuring colour has been a great challenge. Over the past 50 years, biologists have developed and adapted techniques for the objective measurement of colouration that are now widespread. These include spectrometry, photography, and visual modelling.

The application of spectrometry to questions of biological colouration revolutionized the field [@dyck_determination_1966]. Before its widespread use, researchers assigned categorical colour descriptions, arbitrary rankings or scores [e.g. @aldrich_ecogeographic_1991; @bailey_latitudinal_1978; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g. @burtt_analysis_1986; @key_temperature_1954]. Although no doubt applied carefully, these techniques limited progress. For example, many researchers designed colour assessment methods specifically for their projects, rendering data incomparable across studies [e.g. @aldrich_ecogeographic_1991; @bailey_latitudinal_1978]. This is especially problematic when different measurement techniques are used on the same species [e.g. house finch _Carpodacus mexicanus_; @aldrich_origin_1978; @hill_female_1990]. Likewise, because human vision is inherently different to most other animals [@bennett_sexual_1994; @douglas_spectral_2014], its use has lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].

@@

Complementary to spectrometry, digital photography is increasingly being used to quantify biological colouration [@stevens_using_2007]. High-resolution cameras are inexpensive,  and allow instantaneous, repeatable sampling of multiple colour patches [@mckay_use_2013]. The main challenge for gathering colour data from photographs involve ensuring standardization, such as linearization and consistent lighting [@stevens_using_2007]. Also, most digital cameras used for the measurement of biological colouration are consumer products designed for the human visual system. Thus hardware modifications for ultraviolet and infrared imaging are generally necessary [e.g. @stevens_colour_2013].

Expansion in the availability of objective methods for the measurement of colour has been matched by advances in theory and analysis. In particular, the development of sensory models has produced significant insights into how colourful stimuli are perceived [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998]. Sensory models typically attempt to describe the subjective perception of colour or brightness information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g. @endler_comparing_2005; @vorobyev_receptor_1998]. These models enable researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to formulate and test hypotheses.

Although relatively easy to implement, and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and perceptually processed and these can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012]. For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; -@vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e. raw versus log-transformed), the use of any colour constancy mechanism (e.g. von Kries), as well as the photoreceptor signal-to-noise ratios, their relative densities, and the type of noise predominating (i.e. neural and/or receptor noise). These data are only available for a small number of species, such as humans and honeybees, and so assumptions and extrapolations are necessary when models are applied to other species. Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies [@pike_preserving_2012].

 The potential for subjectivity in colour measurement and the mathematical opacity of visual models for the average biologist means that, as in any field, is it critical that methods are reported rigorously. Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011; @sandve_ten_2013]. The aims of this paper were to (a) review methods reporting in papers published on colour in 2013 from 22 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods. Our review specifically focuses on papers that quantify the external coloration of a plant or animal. These may be purely descriptive quantifications of color, or they may be incorporated into models of animal vision. Our hope is that this paper will provide a useful set of guidelines to improve the reproducibility and transparency of studies on colour.

@@

We developed a list of information we deemed critical to be reported for reproducibility of colour methods (Table 1) and analysis (Table 2), loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006]. These included quantitative methods including spectrometric and photographic techniques [as in @kemp_ornamental_2008; @chiao_visualization_2009] but not qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009]. For visual modeling techniques we characterized a set of criteria required for reproducing the analysis (Table 2). Much of these criteria specifically targeted the necessary parameters required to be explicitly defined by the user during analysis. These are essential if such analyses are to be reproduced. These criteria provided that baseline against which to hold our results.

@@

To determine the current state of methods reporting in the field we searched for papers from 2013 in 22 leading journals (_American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_). On each of the journals' homepages, we used the Boolean terms 'colour*', 'color*' or 'spectra*' to search the title or abstract. We read the titles and abstracts of each paper and included only those papers that used either a spectrometer or camera to quantify colouration. We excluded papers review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and papers taking microspectrophotometric measurments of retinal absorbance. To review the large number of papers returned from our search, journals were haphazardly divided up between the authors. In order to reduce the risk of observer bias in our assesment, each paper was read and reassessed by three authors. Any discrepancies between assessor scores were discussed and resolved prior to analysis. We then tabulated what details of the methods were reported against our criteria. The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep (TODO: GIVE IT A DOI?). We have kept the papers used in our dataset annoymous as our aim is to make a general statement oabout methods reporting in the field.

Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis. Overall, `r specs+both` studies used a spectrometer to measure colour, while `r cam+both` used a camera. Of these studies, `r both` quantified colour using both a spectrometer and a camera. Reporting of the make and model of the spectrometer and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 2)`% and cameras = `r round(camModel, digits = 2)`%), however, only `r round(CamLight, digits = 2)`%  and `r round(specLight, digits = 2)`% of camera and spectrometer studies reported a light source. In addition, the number of reflectance spectra and pixels averaged were only reported in `r round(specAVG, digits = 2)`% and `r round(camAVG, digits = 2)`% of studies, respectively. Surprisingly, dark standards were only reported in `r round(drk_std, digits = 2)`% of the spectrometer studies, while white standards were reported in `r round(white_stdspec, digits = 2)`%. Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(intTime, digits = 2)`%), or the angle of the optical probe relative to the surface (`r round(angle, digits = 2)`%) or distance (`r round(dist, digits = 2)`%) the probe was held from the measurement surface.

The types of analysis varied across studies. Overall `r analysisType[1]` papers generated colourimetric variables [i.e. 'spectral' measures of hue, saturation, and/or brightness; @montgomerie_analyzing_2006] whereas `r analysisType[4]` used visual models, `r analysisType[2]` used both colourimetrics and visual models, and `r analysisType[3]` used other forms of analysis (e.g. purely statistical analyses, such as principle component analysis on reflectance spectra). Of the studies quantifying colour using spectral reflectance curves `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models and `r analysisTypeSpec[3]` used other forms of analysis. Studies using colourimetric measurements defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 2)`% of the studies.

@@

While some of the percentages reported above seem alarming, it is important to note that around one-third of papers (`r round(ref_stud, digits = 2)`%) made reference to previous work for various methods. This means that our above figures are slightly liberal given that the referenced works may have comprehensively covered some of these criteria. Despite this, two-thirds of the studies did not reference previous work, and were missing important methodological details, suggesting the need for more comprehensive reporting of data collection and analysis methods within such studies.

@@

Our review of recent colour literature highlights some important reporting gaps that will hinder the reproducibility and comparability of studies of biological colouration (Fig. 1). While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly believe that it is to the field's advantage that researchers report all details essential to a study's transparency and reproducibility wherever possible.

While all criteria are important, some deserve special attention. In particular, we found that the types of light sources, numbers of pixels or spectra averaged, and types of standards used (i.e. white and black reference samples) are under-reported, each of which is critical to repeatability and cross-comparison. Similarly, the distance and angle of the optical probe relative to the measurment surface and collector should be standard reporting for all studies using spectrophometry, as variation in these factors can lead to dramatically different results in some cases (REFS). Visual models were regularly used to analyse spectral data, however, there were many important aspects of the analysis that were seldom reported. These included the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of noise used in receptor-noise models. This is problematic because... Also, photoreceptor densities should be reported where possible even if using standard avian photorecptor densities because.... These all have important consequences on downstream analysis and the reproducibility of colour signal quantification [e.g. @kemp_ornamental_2008; @santos_comparison_2007]. We also recommend the presentation of spectral reflectance curves where possible. These figures allow for the rapid assesment of the nature and quality of the data, independent of downstream processing (if the conditions under which they were captured is comprehensively reported. We found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had a figure showing these curves. When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investigated explicitly outlined.

Our objective in this paper was to highlight common oversights in reporting that can influence the reproducibility of colour research and cross-study comparisons. By no means do we wish to say that studies lacking important details are not of significant value, but rather to provide recommendations for future work. It is only through critical appraisal and collegial discussion that we can identify areas for improvement, and constructively contribute to the growth of our field.

We believe that part of the issue at hand is simply that many researchers seldom have clear guidelines to follow when reporting methods, and may overlook outlining important aspects of methodology and analysis. We hope our critical evaluation of recent literature has highlighted a need for more careful description of how colour signals have been measured and analyzed. More importantly, we hope our guidelines for methods reporting helps establish a standard in this area, to continue the advance of an exciting era in colour research.
